Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from biclaR

### 2024-05-07 paper/PaperCEUS/PaperCEUS_rev1.Rmd
imptove map 3

```{r map1, out.width="60%", fig.cap="Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area.", fig.align='center'}
knitr::include_graphics("img/map1.png", error = FALSE)
```

@@

# re-calculating the percentage of trips
CAR_PT_TRANSF_mun_or_geo_18 = CAR_PT_TRANSF_mun_or_geo_18 |> 
  left_join(TRIPSall_mun, by = c("Concelho" = "Origem"))

ENMAC4 = 0.04 # 4%
ENMAC10 = 0.10 # 10#
CAR_PT_TRANSF_mun_or_geo_18 = CAR_PT_TRANSF_mun_or_geo_18 |> 
  mutate(Bikeper = Bike / viagens,
         Carper = Car / viagens) |>  #should it be "viagens" or "Total  -- all trips, not only the found ones
  mutate(Bike_new4 = ifelse(Bikeper >= ENMAC4 | Carper <= ENMAC4, Bike, ENMAC4 * viagens),
         Bike_new10 = ifelse(Bikeper >= ENMAC10 | Carper <= ENMAC4, Bike, ENMAC10 * viagens)) |> 
  mutate(Bike4 = ifelse(Bikeper >= ENMAC4 | Carper <= ENMAC4, Bike, Bike + Bike_new4),
         Bike10 = ifelse(Bikeper >= ENMAC10 | Carper <= ENMAC4, Bike, Bike + Bike_new10),
         Car_4 = ifelse(Bikeper >= ENMAC4 | Carper <= ENMAC4, Car, Car - Bike_new4),
         Car_10 = ifelse(Bikeper >= ENMAC10 | Carper <= ENMAC4, Car, Car - Bike_new10),
         TP_4 = TP + Bike_new4,
         TP_10 = TP + Bike_new10) |> 
  mutate(Bike_new4per = 100*Bike_new4 / viagens,
         Bike_new10per = 100*Bike_new10 / viagens,
         TPper = 100*TP / viagens,
         TP_4per = 100*TP_4 / viagens,
         TP_10per = 100*TP_10 / viagens)

  #CONTINUE FROM HERE ---  LOOK AT DIFRERENCES TOTAL AND VIAGENS

@@

# plot 3 - the added % of potential bike + TP trips - doesn't make sense to show
ggplot(data = CAR_PT_TRANSF_mun_or_geo_18) +
  geom_sf(aes(fill = Bike_new10per)) + #???
  viridis::scale_fill_viridis(option = "G", direction = -1, name = "new Bike + PT trips %", na.value = "white") +
  theme_void()

>>>>

```{r map1, out.width="60%", fig.cap="Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area.", fig.align='center'}
# knitr::include_graphics("img/map1.png", error = FALSE) #old version

MUNICIPIOSgeo = readRDS("load/MUNICIPIOSgeo.Rds")
stopsAML = sf::st_read("load/Stops_AML.geojson", quiet = TRUE)
stopsAML = stopsAML |> filter(mode != "Metropolitano") |>
  mutate(mode = factor(mode, levels = c("Comboio", "Ferry", "Metro Superficie", "Autocarro"),
                       labels = c("Train", "Ferry", "Tram", "Bus")))
routesAML = sf::st_read("load/Routes_AML.geojson", quiet = TRUE)
routesAML = routesAML |> 
  mutate(mode = factor(mode, levels = c("Comboio", "Ferry", "Metro Superficie", "Autocarro"),
                       labels = c("Train", "Ferry", "Tram", "Bus")))

ggplot() +
  geom_sf(data = stopsAML, aes(color = mode), size = 2) +
  geom_sf(data = routesAML, aes(color = mode), size = 1) +
  scale_color_manual(values = c("purple", "darkgreen", "orange", "dodgerblue")) +
  geom_sf(data = MUNICIPIOSgeo, fill = NA, color = "gray60") +
  theme_void() +
  labs(color = "Mode")
```

@@

# re-calculating the percentage of trips
CAR_PT_TRANSF_mun_or_geo_18 = CAR_PT_TRANSF_mun_or_geo_18 |> 
  left_join(TRIPSall_mun, by = c("Concelho" = "Origem")) |> 
  select(Concelho, Total, Car, Bike, Walk, TP, viagens, geometry)

ENMAC4 = 0.04 # 4%
ENMAC10 = 0.10 # 10#
CAR_PT_TRANSF_mun_or_geo_18 = CAR_PT_TRANSF_mun_or_geo_18 |> 
  mutate(Bikeper = Bike / viagens,
         Carper = Car / viagens) |>  #should it be "viagens" or "Total  -- all trips, not only the found ones
  mutate(Bike_new4 = ifelse(Bikeper <= ENMAC4 & Carper >= ENMAC4, ENMAC4 * viagens, ifelse(Carper <= ENMAC4, Car, Bike)),
         Bike_new10 = ifelse(Bikeper <= ENMAC10 & Carper >= ENMAC10, ENMAC10 * viagens, ifelse(Carper <= ENMAC10, Car, Bike))) |> 
  mutate(Bike4 = Bike +Bike_new4,
         Bike10 = Bike + Bike_new10,
         Car_4 = Car - Bike_new4,
         Car_10 = Car - Bike_new10,
         TP_4 = TP + Bike_new4,
         TP_10 = TP + Bike_new10) |> 
  mutate(Bike_new4per = 100*Bike_new4 / viagens,
         Bike_new10per = 100*Bike_new10 / viagens,
         Bike_4per = 100*Bike4 / viagens,
         Bike_10per = 100*Bike10 / viagens,
         TPper = 100*TP / viagens,
         TP_4per = 100*TP_4 / viagens,
         TP_10per = 100*TP_10 / viagens)

# other way  
CAR_PT_TRANSF_mun_or_geo_18_extrap = CAR_PT_TRANSF_mun_or_geo_18 |> 
  mutate(Bikeper = Bike / Total,
         Carper = Car / Total) |>  #extrapolating to the total trips
  mutate(Bike_new4 = ifelse(Bikeper <= ENMAC4 & Carper >= ENMAC4, ENMAC4 * Total, ifelse(Carper <= ENMAC4, Car, Bike)),
         Bike_new10 = ifelse(Bikeper <= ENMAC10 & Carper >= ENMAC10, ENMAC10 * Total, ifelse(Carper <= ENMAC10, Car, Bike))) |> 
  mutate(Bike4 = Bike +Bike_new4,
         Bike10 = Bike + Bike_new10,
         Car_4 = Car - Bike_new4,
         Car_10 = Car - Bike_new10,
         TP_4 = TP + Bike_new4,
         TP_10 = TP + Bike_new10) |> 
  mutate(Bike_new4per = 100*Bike_new4 / Total,
         Bike_new10per = 100*Bike_new10 / Total,
         Bike_4per = 100*Bike4 / Total,
         Bike_10per = 100*Bike10 / Total,
         TPper = 100*TP / Total,
         TP_4per = 100*TP_4 / Total,
         TP_10per = 100*TP_10 / Total)

@@

# plot 3 - the added % of potential bike + TP trips - doesn't make sense to show
ggplot(data = CAR_PT_TRANSF_mun_or_geo_18) +
  geom_sf(aes(fill = TP_10per- TPper)) + #???
  viridis::scale_fill_viridis(option = "G", direction = -1, name = "new Bike + PT trips %", na.value = "white") +
  theme_void()
